---
title: "DATA605 Homework1"
author: "Jonathan Hernandez"
date: "August 28, 2018"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

# Homework Assignment 1

## Problem Set 1

1. Calculate the dot product u.v where u = [0.5;0.5] and v = [3; -4]

Answer: by using %*% operator to do dot product
```{r question1}
u = c(0.5, 0.5)
v = c(3,-4)

dot_product <- u %*% v
dot_product
```

2. What are the lengths of u and v? Please note that the mathematical notion of 

the length of a vector is not the same as the computer science definition.

Answer:
```{r question2}
# math wise finidng the magnitude of a vector
mag_u <- sqrt(u[1]^2 + u[2]^2)
mag_v <- sqrt(v[1]^2 + v[2]^2)

c(mag_u, mag_v)
```

3. What is the linear combination: 3u - 2v?

Answer:
```{r question3}
3*u - 2*v
```

4. What is the angle between u and v?

Answer: The angle $\theta$ between them is defined as

$\theta$ = $arccos(u*v / (||u|| * ||v||))$ (arc cosine of the inner product of

the two vectors divided by the product of their norms)

```{r question4}
angle <- acos(dot_product / (mag_v * mag_u)) * (180 / pi) # in degrees
angle
```

## Problem Set 2

Create a function that solves a system of equations (3 variables and 3 constraints)

```{r solveSLE}
solveSLE <- function(A, b){
  # get dimensions of matrix A
  m <- cbind(A,b)
  m_row <- nrow(m)
  n_col <- ncol(m)
  pivot <- 1
  
  for(r in 1:m_row) {
    if ( n_col <= pivot ) break; # if only one column or less, exit
    i <- r
    while( m[i,pivot] == 0 ) { # for entries for 0 
      i <- i + 1 # go to next row
      if ( m_row == i ) { # find pivot
        i <- r
        pivot <- pivot + 1 # move pivot until m[i,pivot] is not zero
        if ( n_col == pivot ) return(m)
      }
    }
    # swap rows (next three lines)
    trow <- m[i, ]
    m[i, ] <- m[r, ]
    m[r, ] <- trow
    m[r, ] <- m[r, ] / m[r, pivot]
    for(i in 1:m_row) {
      if ( i != r )
        m[i, ] <- m[i, ] - m[r, ] * m[i, pivot] # row operations
    }
    pivot <- pivot + 1
  }
  return(m[,n_col]) # last column is the x coefficients
}

# test out function
A <- matrix(nrow = 3, ncol = 3, data = c(1, 2, -1, 1, -1, -2, 3, 5, 4))
b <- c(1, 2, 6)
solveSLE(A, b)
```