---
title: "Homework 11"
author: "Jonathan Hernandez"
date: "November 2, 2018"
output: pdf_document
---

    Using the "cars" dataset in R, build a linear model for stopping distance as a function of speed and
    replicate the analysis of your textbook chapter 3 (visualization, quality evaluation of the model, and
    residual analysis.)


### Load the built-in R "cars" dataset

```{r}
data("cars")
head(cars, n = 10)
```

### Summary of the dataset

```{r}
summary(cars)
```

### Dimensions of the dataset

```{r}
dim(cars)
```

### Visualize the data

```{r}
with(cars, plot(speed, dist, 
                xlab = "Speed (mph)",
                ylab = "Stopping Distance (ft)"))
```

### Fit a Linear Model $distance = a_0 + a_1*speed$

```{r}
# linear model
lm_cars <- lm(dist ~ speed, data = cars)
lm_cars
```

- Thus y-intercept or $a_0 = -17.579$ and the slope $a_1 = 3.932$
and the linear model is

```{r}
with(cars, plot(speed, dist, 
                xlab = "Speed (mph)",
                ylab = "Stopping Distance (ft)"))
abline(lm_cars)
```

$$
\begin{aligned}
dist = -17.579 + 3.932*speed
\end{aligned}
$$

### Quality of the Model

```{r}
summary(lm_cars)
```

- We can see the summary statistics of the residuals which are the differences between

the actual measured values and the values on the line. A good-fit model would have the residuals

to be nearly standard normal. The median should be near 0 which is the case here.

```{r}
hist(lm_cars$residuals, xlab = "", main = "Histogram of Residuals")
```

- A key statistic is $R^2$ value which shows that the model explains about 65% of the

data's variation which for a linear model is not too bad.

- Also looking at the p-values, we see that the probability that the speed variable is not

relevant is very small at about $1.49 * 10^{-12}$ This means that speed plays a key predictor in

determining stopping distance and a strong dependency.

### Residual Analysis

```{r}
plot(fitted(lm_cars), resid(lm_cars))
```

- Residuals are nearly uniformly scattered and approximately constant variance.

- A Quantile vs Quantile or Q-Q plot
```{r}
qqnorm(resid(lm_cars))
qqline(resid(lm_cars))
```