---
title: "Week 5 Assignment - Tidying and Transforming Data"
author: "Jonathan Hernandez"
date: "September 29, 2016"
output: html_document
---

#### This assignment is to show how given a "wide" dataset structure using dplyr

#### and tidyr, make the dataset clean and tidy in this case examing arrival

#### delays for two airlines across five destinations and then perform analysis

#### on the tidy dataset.

#### Goal is to get the dataset to have the following format

#### airline  destination   arrival

```{r getflights, echo = TRUE}
library(dplyr)
library(tidyr)
library(magrittr) # for using the pipe operators for dplyr/tidyr

arrflights <- read.csv("/home/jonboy1987/Desktop/CUNYSPS/IS607/Assignments/Week5/flight_arrivals.csv")
arrflights

# Change some column names to be more readable
colnames(arrflights)[1:2] <- c("airline","arrival")
names(arrflights)
```

#### Lets make the data more tidy and better formated

```{r tidyarrflights, echo = TRUE}
# strip off the 3rd row as it really has no data associated with it
arrflights <- arrflights[complete.cases(arrflights), ]

# fill in the gaps for the airline so two for Alaska, two for AM Flights
arrflights$airline <- c(rep("ALASKA", 2), rep("AM WEST", 2))

# gather the number of flights by destination and then arrange the data to be
# sorted by airline and then by destination
tidy_arrflights <- arrflights %>% gather(destination, flights, 3:7) %>% 
  arrange(airline, destination)

# replace the '.' in the destination with ' '
tidy_arrflights$destination <- gsub("\\.", " ", tidy_arrflights$destination)

tidy_arrflights <- tidy_arrflights %>% mutate_each(funs(factor), -flights)
tidy_arrflights
```

#### With our data clean, now we can perform some EDA (Exploratory Data Analysis)

```{r EDA_tidy_arrflights, echo = TRUE}
str(tidy_arrflights)

# What kind of distribution does destination ~ log(flights) show???
par(cex.axis = .75)
with(tidy_arrflights, plot(destination, log(flights),
                           main = "destination vs number of flights",
                           xlab = "destination", ylab = "log(flights)"))

# and the same for airlines vs flights?
with(tidy_arrflights, plot(airline, log(flights),
                           main = "airline vs number of flights",
                           xlab = "airline", ylab = "log(flights)"))
```

#### It seems like the flights have roughly the same mean and median when 

#### Compared with different airlines and destinations and possible normally

#### distributed.