---
title: "Lecture 2 - Tax lot data NYC"
author: "Jonathan Heranndez"
date: "September 13, 2017"
output: html_document
---

```{r test, echo= FALSE, include = FALSE}
library(dplyr)
library(ggplot2)
source("all_PLUTO_data.R")
```

```{r combineboroughs, echo = TRUE, cache = TRUE}
# combine all 5 boroughs dataset into one..
nycPluto <- rbind(BK,BX,QN,SI,MN, colNames = FALSE, fill = TRUE)
nycPluto <- nycPluto[, -c("x")] # get rid of extra column formed by above; not needed
```

```{r question1, echo = TRUE}
# get buildings that were built after 1850 and later
nycPluto1850 <- nycPluto %>% filter(YearBuilt > 1850) %>% 
  select(YearBuilt, NumBldgs)

# ggplot but use the log scale for y to better see the number of buildings per year
ggplot(data = nycPluto1850, aes(x = YearBuilt, y = NumBldgs)) +
  geom_bar(stat = "summary", fun.y = "log10", na.rm = TRUE) +
  ggtitle("NYC Buildings built since 1850") +
  ylab("log(Number of Buildings Built)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.border = element_blank(), panel.background = element_blank(),
        axis.title.x = element_blank())
```

#### Looking at the data many buildings were built around the 1920's to 1940's

#### roughly. Seeing very few buildings based on the data makes sense as that was 

#### during the industrial revolution and apartment buildings started being built slowly.

#### seeing few buildings especially after around 1960/1970 also makes sense as many

#### nyc buildings are mostly pre-war buildings and few buildings and space for post-war buildings.

```{r question2, echo = TRUE}
# tall buildings at least 20 floors or more
tallNYCBuildings <- nycPluto %>% filter(NumFloors >= 20 & YearBuilt > 1850) %>%
  select(NumFloors, NumBldgs, YearBuilt)

# see the unique number of floors
x <- tallNYCBuildings$NumFloors

# split number of Floors into ranges for graphing (buildings with 20-30, 30-40,...)
floor_ranges <- cut(x, seq(from = 19.9, to = 120, by = 10),
                    labels = c("20-story", "30-story", "40-story", "50-story", 
                               "60-story", "70-story", "80-story", "90-story",
                               "100-story", "110-story"))

# now plot number of buildings each year but add a facet to have multiple graphs/grids by floor range
ggplot(data = tallNYCBuildings, aes(x = YearBuilt, y = NumBldgs)) +
  geom_bar(stat = "summary", fun.y = "log10", na.rm = TRUE) +
  facet_grid(as.factor(floor_ranges) ~.) +
  ggtitle("NYC Buildings built since 1850") +
  ylab("log(Number of Buildings Built)") +
  theme_bw() +
  theme(panel.border = element_blank(), panel.background = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 7, angle = 45, hjust = 1, vjust = 1.25)) +
  ggsave(file = "Hw2_figure2.png")
```