---
title: "Character Manipulation and Data Processing"
author: "Jonathan Hernandez"
date: "September 10, 2016"
output: html_document
---

```{r question3, echo=TRUE}
name <- c("Moe Szyslak", "Burns, C. Montgomery", "Rev. Timothy Lovejoy",
          "Ned Flanders", "Simpson, Homer", "Dr. Julius Hibbert")

# 3. Rearrange the vector so that all elements conform to the standard
# first_name last_name

# split the vector based on spaces
name_split <- strsplit(name, " ")
name_split

check_name_prefix <- function(nameString)
{
   regexpr("\\.", nameString[1])
}

sapply(name, check_name_prefix)

```

### 1. The regexular expression "[0-9]+\\$" matches any string that ends with

### one or more digits followed by a \ character

### 2. The regular expression "\\b[a-z]{1,4}\\b" matches any string that starts

### with a \b followed by 1-4 characters of the alphabet followed by a \b

### 3. The regular expression ".*?\\.txt$" matches any string that ends with a 

### \.txt at the end

### 4. The regular expression "\\d{2}/\\d{2}/\\d{4}" matches only the following

### string \dd\dd\dddd

### 5. The regular expression ""<(.+?)>.+?</\\1>" matches a string starting with

### one or more characters enclosed in angular brackets then followed by one or

### more characters followed by <\1> at the end.

```{r question4, echo = TRUE}
prob4.1 <- "[0-9]+\\$"

example_prob4.1 <- regexpr(prob4.1, "mypassword2016\\")
example_prob4.1

prob4.2 <- "\\b[a-z]{1,4}\\b"
# Examples:
# "\bobby\b"
# "\bush\b"
# "\boy\b"

example_prob4.2 <- regexpr(prob4.2, "\bobby\b")
example_prob4.2

prob4.3 <- ".*?\\.txt$"
# Examples:
# "CUNYSPS\.txt"
# "nc9aqt\.txt"

example_prob4.3 <- regexpr(prob4.3, "mnvac4534\\.txt")
example_prob4.3

prob4.4 <- "\\d{2}/\\d{2}\\d{4}"
example_prob4.4 <- regexpr(prob4.4, "\\dd\\dd\\dddd")
example_prob4.4

prob4.5 <- "<(.+?)>.+?</\\1>"
example_prob4.5 <- regexpr(prob4.5, "<nfoaetyey49y5>j437n9m8byt<\1>")
example_prob4.5
```