---
title: "Week 11 Regression Analysis in R"
author: "Jonathan Hernandez"
date: "November 3, 2018"
output:
  html_document: default
---

    Using R, build a  regression model for data that interests you.  Conduct residual analysis.
    Was the linear model appropriate? Why or why not?

### For this discussion, I will look at Kaggle's Powerlifting Database dataset.

- It's a dataset containing competitor results in powerlifting from the 

OpenPowerlifting Database and do residual analysis.

I will build a simple linear regression model of body weights vs best bench press for seniors to see if 

a linear relation exists between them.

- Dataset can be found here: https://www.kaggle.com/open-powerlifting/powerlifting-database

- Overview: https://www.kaggle.com/open-powerlifting/powerlifting-database/home

- Get the data and examine a preview

```{r}
library(data.table)
powerlift <- fread("openpowerlifting.csv")
head(powerlift, n=5)
```

- Summary of the data
```{r}
summary(powerlift)
```

- Dimensions of the data and its structure
```{r}
dim(powerlift)
str(powerlift)
```

- Grab rows such that the Best Bench Press > 0 (those competitors who were most likley
disqualified) and Division as Seniors and extract just the 2 attributes in question

```{r, message=FALSE}
library(dplyr)
powerlift_senior <- powerlift %>% filter(BestBenchKg > 0 & Division %like% 'Senior') %>% 
  select(BodyweightKg, BestBenchKg)
```

### Visualize the data (EDA)

```{r, message=FALSE}
with(powerlift_senior, plot(BodyweightKg, BestBenchKg, xlab = "Body Weight (Kg)",
                            ylab = "Best Bench Press (Kg)"))
```

### Residual Analysis

```{r}
lm_powerlift_senior <- lm(BestBenchKg ~ BodyweightKg, data = powerlift_senior)
with(powerlift_senior, plot(BodyweightKg, BestBenchKg,xlab = "Body Weight (Kg)",
                            ylab = "Best Bench Press (Kg)"))
abline(lm_powerlift_senior)
summary(lm_powerlift_senior)
```

- Equation of line is

$$
\begin{aligned}
\widehat{bestbenchpress} = 0.1376 + 1.2804*\widehat{bodyweight}
\end{aligned}
$$

- Residual plots
```{r}
plot(fitted(lm_powerlift_senior), resid(lm_powerlift_senior),
     main = "Fitted vs residuals", xlab = "", ylab = "")
abline(h =0)
hist(resid(lm_powerlift_senior), xlab = "", main = "Histogram of Residuals")
qqnorm(resid(lm_powerlift_senior))
qqline(resid(lm_powerlift_senior))
```

### Summary

- We see that a linear model based on one explanatory variable doesn't explain the

data well. The $R^2$ value is quite low which shows that the fitted model doesn't

accuractely predict the values of Senior divisions competitiors bench press best based 

on their weight.

- Looking at the summary of the model also shows that the intercept has a high p-value

meaning that the significance can be ignored and is not a good estimate. One good thing

is that the p-value of estimated body weight is a good predictor due to the low p-value

and standard error.

- The residual plots show that even though the distribution of residuals is nearly

normal, the fitted data vs the residuals doesn't show constant variance and for 

larger fitted values, the variance begins to change. A good-fit linear model would 

have constant variance or nearly constant variance.