---
title: "Examing Basball Players data Pre and Post PEDs"
author: "Jonathan Hernandez"
date: "November 19, 2016"
output: html_document
---

## This Project will look at the Lahman database on MLB statistics and from the 

## Wikipedia list of players who used Performance enhancing Drugs and compare the

## stats for them before announcement and post announcement of use of PED's.

## While many will think that the performance of Athelics will increase by taking

## PED's, it will be interesting to see if that is the case or not.

```{r loadLibraries, echo = FALSE, include = FALSE}
if (!require(RMySQL)) install.packages("RMySQL")
if (!require(rvest)) install.packages("rvest")
if (!require(ggplot2)) install.packages("ggplot2")
if (!require(dplyr)) install.packages("dplyr")
if (!require(stringr)) install.packages("stringr")
```

#### Get the data (wikipedia site)
```{r getPEDUsers, echo = TRUE, cache = TRUE}
bb_player_PED <- read.csv("Baseball_Players_PED.csv")
head(bb_player_PED, 10)
```

#### Go into the Lahman Databse 

#### (http://www.seanlahman.com/baseball-archive/statistics/)

#### (you have three options, donwload it as a csv, MS Access or SQL) for the

#### challenge I decided to go with SQL instead of being used to csv files

#### There are archive data that contains previous MLB season but the one I will

#### use is 2015 MLB (at the time of this project the latest is 2015 as that is

#### the most recent one)

```{r getLahaman_bb_Data, echo = TRUE, cache = TRUE}
# connect to database
con <- dbConnect(MySQL(), user = "cunysps", password = "cunyspsdb",
                 dbname = "stats", host = "127.0.0.1")

# Query to get baseball battings statistics during regular season since 1990
bb_batting_query <- "select b.*, m.nameFirst, m.nameLast, m.nameGiven, m.bats, m.throws, m.debut, m.finalGame
from batting b inner join stats.master m
on b.playerID = m.playerID
where b.yearID >= 1990;"

baseball_batting_data <- dbGetQuery(con, bb_batting_query)
summary(baseball_batting_data)

# Query to get baseball pitching statistics during regular season since 1990
bb_pitching_query <- "select p.*, m.nameFirst, m.nameLast, m.nameGiven, m.bats, m.throws, m.debut, m.finalGame
from pitching p inner join stats.master m
on p.playerID = m.playerID
where p.yearID >= 1990;"

baseball_pitching_data <- dbGetQuery(con, bb_pitching_query)
summary(baseball_pitching_data)

# get teamID and team names throughout since 1990
bb_teams_query <- "select distinct t.yearID, t.name, t.teamID, t.lgID
from stats.teams t where t.yearID >= 1990;"

baseball_teams_data <- dbGetQuery(con, bb_teams_query)
summary(baseball_teams_data)
dbDisconnect(conn = con)
```

#### Begin Cleaning the dataset(s)

```{r clean_PED_data, echo = TRUE}
# remove some columns from the baseball_players_PED list
# remove response and reference columns
bb_player_PED_clean <- bb_player_PED %>% select(-Ref.) %>% select(-Response)

# change the "Date.Announced" name to something more meaningful like "Date_Caught"
colnames(bb_player_PED_clean)[3] <- "Date_Caught"

# change the data type of Date_Caught to be of character format
bb_player_PED_clean$Date_Caught <- as.character(bb_player_PED_clean$Date_Caught)
# in the Penalty field, remove any '^' characters, [A] or [B] or [C] string
bb_player_PED_clean$Penalty <- gsub("(\\[A\\]|\\^|\\[B\\]|\\[C\\])", ""
                              ,bb_player_PED_clean$Penalty)

# in the Player's Name field, remove any "([0-9])" characters 
# and ("SS, GG, MVP") as it is not relvalent for this data analysis if a player
# was a Golden Glove or voted most valuable player.
bb_player_PED_clean$Player <- gsub("(SS|GG|MVP)", ""
                                   ,bb_player_PED_clean$Player)

bb_player_PED_clean$Player <- gsub("(\\([0-9]\\))", ""
                                   ,bb_player_PED_clean$Player)

# Trim white spaces
bb_player_PED_clean$Player <- str_trim(bb_player_PED_clean$Player)

# Remove accents in baseball player names (this will help us in cross-ref names
# with the lahman database)
bb_player_PED_clean$Player <- iconv(bb_player_PED_clean$Player,
                                    to = 'ASCII//TRANSLIT')
bb_player_PED_clean
```

#### Exploratory Data Analysis

#### Here we will compare some player stats to get some insight

```{r baseball_EDA, echo = TRUE}
# graph baseball_ped user stats for a random set of people
# get random rows
PED_sample_rows <- sample(1:nrow(bb_player_PED_clean), 10 , replace = TRUE)

# get data indexed by those random rows
sample_baseball <- bb_player_PED_clean %>% slice(PED_sample_rows)

for (i in 1:nrow(sample_baseball)) {
  # get the year the baseball player was suspended and caught
  year <- regmatches(sample_baseball$Date_Caught[i],
                     regexpr("([0-9]){4}", sample_baseball$Date_Caught[i]))
  # plot the data (if it is a pitcher, display stats such as ERA, SO and BB)
    # split the Player name into first and last name to filter the data from
    # the Lahman database
    fname <- with(sample_baseball, head((str_split(Player[i], " "))[[1]], n = 1))
    lname <- with(sample_baseball, tail((str_split(Player[i], " "))[[1]], n = 1))
    if (sample_baseball$Position[i] == "P"){
      tmp <- baseball_pitching_data %>% 
        filter(grepl(fname, nameFirst),
               nameLast == lname)
      print(tmp)
      print(ggplot(data = tmp, aes(yearID, ERA)) + 
              geom_point(aes(color = "cyl")) + 
              geom_line(aes(color = "blue")) + 
              geom_vline(xintercept = as.integer(year)))
    }
    else {
       tmp <- baseball_batting_data %>% 
        filter(grepl(fname, nameFirst),
               nameLast == lname) 
       print(tmp)
       
       print(ggplot(data = tmp, aes(yearID, HR)) + 
               geom_point(aes(color = "cyl")) + 
               geom_line(aes(color = "blue")) + 
               geom_vline(xintercept = as.integer(year)))
    }
}
```