---
title: "Homework 10"
author: "Jonathan Hernandez"
date: "November 2, 2018"
output: pdf_document
---

    Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars.
    A guard agrees to make a series of bets with him. If Smith bets A dollars, he
    wins A dollars with probability 0.4 and loses A dollars with probability 0.6.
    
    Find the probability that he wins 8 dollars before losing all his money if
    
    (a) he bets 1 dollar each time (timid strategy)
    (b) he bets, each time, as much as possible but not more than necessary to
    bring his fortune up to 8 dollars (bold strategy).
    (c) Which strategy gives Smith the better chance of getting out of jail?
    
let $S = \{s_1, s_2,...,s_8\}$ be the states where $s_i =$ state of having i dollars.

let p = probability he wins A dollars and q = 1-p = probability he loses A dollars

for (a) with this given information we have the transition matrix

```{r}
# transition matrix
p <- 0.4
q <- 0.6
P <- matrix(c(1, 0, 0, 0, 0, 0, 0, 0, 0,
              q, 0, p, 0, 0, 0, 0, 0, 0,
              0, q, 0, p, 0, 0, 0, 0, 0,
              0, 0, q, 0, p, 0, 0, 0, 0,
              0, 0, 0, q, 0, p, 0, 0, 0,
              0, 0, 0, 0, q, 0, p, 0, 0,
              0, 0, 0, 0, 0, q, 0, p, 0,
              0, 0, 0, 0, 0, 0, q, 0, p,
              0, 0, 0, 0, 0, 0, 0, 0, 1), nrow = 9, ncol = 9, byrow = TRUE)
rownames(P) <- 0:8
colnames(P) <- 0:8
P
```

We can compute the probability of reaching the absorbing state that is at state $s_8$

by computing $B = NR$ where $N = (I-Q)^{-1}$ and $I =$ 7x7 identity matrix and 

$Q =$ square matrix of the transient states that is rows and columns 1:7 and

$R=$ square matrix of the column absorbing states and the row transient states

that is

```{r}
library(Matrix)
# matricies to build matrix B as the probability
Q <- P[c(2:8), c(2:8)]
I <- diag(nrow = nrow(Q))
Q

N <- solve(I - Q)
N

R <- P[c(2:8), c(1, 9)]
R

# compute B

B <- N %*% R
B
round(B[1,2], 4)
```

By looking at the first row second column of B which represents the probability that at the 

start of this series of bets, he has only about 0.0203 or about 2.03% percent chance of getting

out of jail.

for (b) the transition matrix will be different and smith will go for a all-or-nothing approach.

the transition matrix will look something like this and re-calculation of the matricies:

```{r}
P <- matrix(c(1, 0, 0, 0, 0, 0, 0, 0, 0,
              q, 0, p, 0, 0, 0, 0, 0, 0,
              q, 0, 0, 0, p, 0, 0, 0, 0,
              0, 0, 0, 0, 0, 0, 0, 0, 0,
              q, 0, 0, 0, 0, 0, 0, 0, p,
              0, 0, 0, 0, 0, 0, 0, 0, 0,
              0, 0, 0, 0, 0, 0, 0, 0, 0,
              0, 0, 0, 0, 0, 0, 0, 0, 0,
              0, 0, 0, 0, 0, 0, 0, 0, 1), nrow = 9, ncol = 9, byrow = TRUE)
rownames(P) <- 0:8
colnames(P) <- 0:8
P

Q <- P[c(2:8), c(2:8)]
I <- diag(nrow = nrow(Q))
Q

N <- solve(I - Q)
N

R <- P[c(2:8), c(1, 9)]
R

# compute B

B <- N %*% R
B
round(B[1,2], 4)
```

We see that using this all-or-nothing approach will give Smith about 6.4% of getting

out of jail.

for (c), we see that going the bold route/all-or-nothing approach is more feasible than

the timid approach and more likley for Smith to get out of jail.