---
title: "Homework4"
author: "Jonathan Hernandez"
date: "September 13, 2018"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

## Homework Assignment 4

## Problem set 1

```{r}
A <- matrix(c(1, -1, 2, 0, 3, 4), nrow = 2, ncol = 3)
A

# compute X = AA^T
X <- A %*% t(A)

# compute Y = A^TA
Y <- t(A) %*% A

print("Matricies X and Y: ")
list(X,Y)

# Compute eigenvalues and eigenvectors of X and Y
print("Eigenvalues and Eigenvectors of X and Y:")
eig_X <- eigen(X)
eig_Y <- eigen(Y)

list(eig_X, eig_Y)

# compute SVD 

svd_A <- svd(A, nu = 2, nv = 3)
svd_A
```

### Let's look at the singular vectors d, left singular u and right singular v

### for A = UDV' where V' is the complex conjugate (or in this case of real numbers)

### just the conjugate

### For X an d Y

```{r}
list(eig_X$vectors, svd_A$u)

list(eig_Y$vectors, svd_A$v)
```

### We can see that the values (not the sign) for u and v are the eigenvectors

### of X and Y respectively.

### Let's look at the non-zero eigen values of X and Y and show they are the same:

```{r}
list(eig_X$values, eig_Y$values[1:2])
```

### And squaring of either the eigenvalue of X or Y yields the non-zero singular values of A

```{r}
list(eig_X$values, svd_A$d[1]^2)

list(eig_Y$values[1:2], svd_A$d[2]^2)
```


```{r}
# create the function myinverse which will be a function that computes the inverse of a 
# well-conditioned full-rank square matrix using co-factors

myinverse <- function(A){
  # check if det(A) = 0 otherwise continue
  n <- nrow(A) # number of rows
  if( det(A) == 0 || n == 1){
    return("Matrix is not invertible. Exiting")
  }
  
  if (n == 2){ # 2x2 matrix compute inverse manually
    return((1 / det(A)) * matrix(c(A[2, 2], -A[2, 1], -A[1, 2], A[1, 1]),
                                 nrow = 2, ncol = 2))
  }
  
  # create cofactor matrix set it to be initially all 0's
  # simple case a 2 x 2 matrix
  C <- matrix(0, nrow = n, ncol = n)
  # loop through whole matrix A computing minors and cofactor matrix entries
  for(i in 1:n){
      for(j in 1:n){
        minor <- det(A[-i, -j]) # determinant of minor[i,j]
        C[i, j] <- (-1)^(i + j) * minor # cofactor matrix C[i,j]
      }
  }
  # with the co-factor matrix complete, the inverse of A that is A^-1 is
  # A^-1 = 1/det(A)*t(C) where t(C) is the transpose of the conjugate matrix
  return((1 / det(A))*t(C))
}
A <- matrix(c(1, 2, 4, 7, 1,
              2, 1, -1, 3, 0,
              4, -1, -2, 6, -1,
              3, 1, 5, 2, 1, 
              0, 3, 1, 1, 1), nrow = 5, ncol = 5)
B <- myinverse(A)
B
```

### Show that by multplying the matrix A with it's inverse A^-1, we get the Identity

### matrix

```{r}
A %*% B
```

### As we can see the off-diagonals are very close to 0 and the diagonal entries

### equal to 1.