---
title: "Spring 2018 DATA606 Final Project Proposal"
author: "Jonathan Hernandez"
date: "April 30, 2018"
output: html_document
---

```{r school-progress-report, echo=FALSE, cache=TRUE, results='hide'}
school_progress_report_scores <- read.csv("2006-2007_School_Progress_Report.csv")
```

This is the DATA606 Project Proposal. This project will look at a historical NYC

School Progress Report dataset from 2006-2007.

Research Question:

Is the overall school rating predictive of various categorical scores as well as

predictive of Borough of the school, grades and school level?

Cases:

There are 

```{r,echo=FALSE}
nrow(school_progress_report_scores)
```
in the dataset.

Type of Study:

This is an observational study on past data.

Data Source:

Dataset is available here: 

https://data.cityofnewyork.us/Education/2006-2007-School-Progress-Report/weg5-33pj

The data comes from the NYC OpenData website provided by the Department of 
Education (DOE).

Explanatory Variables:

The explanatory variables in question are:

- DBN (District Borough Number) - categorical
- SCHOOL LEVEL - categorical 
- GRADE - categorical ordinal
- ENVIRONMENT CATEGORY SCORE - numerical continous
- PERFORMANCE CATEGORY SCORE - numerical continous
- PROGRESS CATEGORY SCORE - numerical continous
- QUALITY REVIEW SCORE - categorical ordinal

Response Variable

The response variable is 

- OVERALL SCORE - numerical continous

Relevant Summary Statistics:

- Understanding the structure and summary
```{r}
str(school_progress_report_scores)
```

Data Cleanup:

Convert score variables to numeric from factor
```{r}
school_progress_report_scores$ENVIRONMENT.CATEGORY.SCORE <- 
  as.numeric(as.character(school_progress_report_scores$ENVIRONMENT.CATEGORY.SCORE))

school_progress_report_scores$PERFORMANCE.CATEGORY.SCORE <- 
  as.numeric(as.character(school_progress_report_scores$PERFORMANCE.CATEGORY.SCORE))

school_progress_report_scores$PROGRESS.CATEGORY.SCORE <- 
  as.numeric(as.character(school_progress_report_scores$PROGRESS.CATEGORY.SCORE))

school_progress_report_scores$OVERALL.SCORE <- 
  as.numeric(as.character(school_progress_report_scores$OVERALL.SCORE))
```
Exploratory Data Analysis:

```{r EDA-hist}
hist(school_progress_report_scores$ENVIRONMENT.CATEGORY.SCORE)
hist(school_progress_report_scores$PERFORMANCE.CATEGORY.SCORE)
hist(school_progress_report_scores$PROGRESS.CATEGORY.SCORE)
```
The 3 histograms of our numerical predictors are very near-normal

```{r EDA-plot-categorical}
with(school_progress_report_scores, plot(OVERALL.SCORE ~ GRADE))
with(school_progress_report_scores, plot(OVERALL.SCORE ~ SCHOOL.LEVEL.))
```
