---
title: "Working with HTML, XML and, JSON Files"
author: "Jonathan Hernandez"
date: "October 14, 2016"
output: html_document
---

#### This Assignment reads 3 files from the web (Github) which are HTML, XML and

#### JSON. The data is 3 books with attributes such as title, author, ISBN,

#### number of pages etc. Lets see how R reads in these three files

```{r loadLibaries}
library(jsonlite)
library(XML)
library(xml2)
library(RCurl)
library(rvest)
library(dplyr)
library(plyr)
```

#### Load JSON data

```{r BooksJSON, echo = TRUE}
jsonURL <- "https://raw.githubusercontent.com/jonathan1987/CUNYSPS_IS607/master/Week7/books.json"
books_json <- fromJSON(jsonURL)
```

```{r BooksHTML, echo = TRUE}
htmlURL <- "https://raw.githubusercontent.com/jonathan1987/CUNYSPS_IS607/master/Week7/books.html"
books_HTML <- read_html(htmlURL)
books_HTML <- html_table(books_HTML)
```

```{r BooksXML, echo = TRUE}
xmlURL <- "https://raw.githubusercontent.com/jonathan1987/CUNYSPS_IS607/master/Week7/Books.xml"
books_XML <- xmlRoot(xmlParse(getURL(xmlURL))) # get XML file contents

# make into a dataset with ldply
books_XML_df <- ldply(xmlToList(books_XML), data.frame)

# remove the .id column
books_XML_df <- books_XML_df %>% select(-.id)
```

#### Print out each dataset

```{r printDatasets, echo = TRUE}

books_json

books_HTML

books_XML_df
```

#### We can see that the datasets are similar except for when it comes to

#### multiple objects (authors in this case)

#### JSON takes all authors and puts it in one column

#### HTML has authors for a book across multiple columns

#### XML like HTML takes multiples authors for a books and puts them in 

#### multiple columns.