---
title: "Cleaning Various Datasets"
author: "Jonathan Hernandez"
date: "October 4, 2016"
output: html_document
---

#### This Project is for analyzing 3 different datasets from forums on the

#### CUNY SPS IS607 Fall 2016, clean and tidy the data using dplyr and tidyr

#### and then analyzing the data.

```{r importLibaries, include = FALSE}
library(dplyr)
library(tidyr)
library(magrittr)
library(data.table)
```

#### First dataset to look at; a table showing income based on various religion

#### traditions.

```{r dataset1, echo = TRUE, cache = TRUE}
# load the income_religion dataset
income_religion <- read.csv("/home/jonboy1987/Desktop/CUNYSPS/IS607/Assignments/Project2/religion_income.csv")
names(income_religion)

# Change the attribute names of the dataset
colnames(income_religion) <- c("Religion", "Low_Class", 
                               "Middle_Class_1", "Middle_Class_2",
                               "Upper_Middle_Class", "Sample_Size")

income_religion
```

#### Data Dictionary

* Low_class --> income < $30,000
* Middle_Class_1 --> income between $30,000 and $49,999 inclusive
* Middle_Class_2 --> income between $50,000 and $99,999 inclusive
* Upper_Middle_Class --> income $100k or more

#### Let's use the dplyr and tidyr packages and combine the incomes into 

#### one column

```{r cleanDataset1, echo = TRUE, warning = FALSE}
# Tidy up the dataset by grouping the income levels into a single column 
tidy_income_religion <- income_religion %>% gather(Working_Class,
                                                   percentage_income, 2:5)

# Check out the new dataset
tidy_income_religion
summary(tidy_income_religion)
```

```{r EDADataset1, echo = TRUE}
# As people make more money, do percentage of Jehova witnesses decrease?
JehovaW <- filter(tidy_income_religion, Religion == "Jehovah's Witness")
JehovaW
```

#### Second dataset to examine (Social Networking Use)

#### http://www.pewresearch.org/data-trend/media-and-technology/social-networking-use/

```{r getDataset2, echo = TRUE, cache = TRUE}
social_net_use <- read.csv('/home/jonboy1987/Desktop/CUNYSPS/IS607/Assignments/Project2/socialNetworkingUse.csv')
social_net_use

# Change the attribute names around
colnames(social_net_use) <- c("Date", "All_Internet_Users", "Young_Adults",
                              "Adults_Middleage1", "Adults_Middleage2",
                              "Seniors")
names(social_net_use)
```

```{r cleanDataset2, echo = TRUE}
# combine the age groups into one variable namely Age for a more tidy dataset

tidy_social_net_use <- social_net_use %>% gather(Age_Group, num_users, 2:6)
tidy_social_net_use
```

#### Data Dictionary for the age groups in the tidy dataset:

* Young_Adults --> people who were 18-29 years of age
* Adults_Middleage1 --> people who were 30-49 years of age
* Adults_Middleage2 --> people who were 50-64 years of age
* Seniors --> people who were 65+ years of age

#### Do some Analysis on the second dataset; A dataset that shows Social Network

#### usage amongst different age groups as of July 2015

```{r EDADataset2, echo = TRUE, warning = FALSE}
# Show the boxplots of each type of age group

# convert Age_group to factors and not characters
tidy_social_net_use <- tidy_social_net_use %>% 
  mutate_each(funs(factor), Age_Group) %>%
  mutate_each(funs(as.numeric), num_users)

# Plot the boxplot
with(tidy_social_net_use,
                 plot(Age_Group, num_users,
                      main = "Boxplot of Internet Users for each age bracket",
                      ylab = "Number of Internet Users", xaxt = "n"))
labels <- tidy_social_net_use$Age_Group
text(labels, labels = labels, 
     par("usr")[3] - .25, srt = 45, xpd = TRUE, adj = 1, cex.axis = .75)
```

#### We can see that the distribution of number of internet users is just about

#### right skewed for age brackets. The median for almost each age group can

#### perhaps indicate that over the years, people began using the internet more.

#### Now to examine the final dataset; a dataset that shows crimes in

#### Chicago, IL since 2001

### Note: this is a ~ 1.4 GB dataset so make sure you have enough RAM to load

### into R.

#### data can be found at https://data.cityofchicago.org/view/5cd6-ry5g

```{r getDataset3, echo = TRUE, cache = TRUE, include = FALSE}
chicago_crimes <- fread('/home/jonboy1987/Desktop/CUNYSPS/IS607/Assignments/Project2/Crimes_-_2001_to_present.csv')
```

#### Examine the dataset in more detail and get a sense of the data

```{r, echo = TRUE}
str(chicago_crimes)
summary(chicago_crimes)
```

#### Examine the crimes that had an arrest and where the primary reason and/or

#### description was either an "assault" or "theft" only

#### But lets clean some portion of this dataset first

```{r cleanDataset3, echo = TRUE}
# update underscores of the attribute names with a '_' symbol so the attribute
# names can be referenced with '$'
colnames(chicago_crimes) <- gsub(" ", "_", colnames(chicago_crimes))

# make the Primary_Type, Description, "Arrest" factor variables instead of 
# character vectors (better to work with)
tidy_chicago_crimes <- chicago_crimes %>%
  mutate_each_(funs(factor), c("Primary_Type", "Description", "Arrest"))

# select the data where Arrest == true and the primary type or description have
# the words assault or theft in them

tidy_chicago_crimes_arrest <- tidy_chicago_crimes %>% filter(Arrest == "true")

# Any arrests dealing with assault?
tidy_chicago_crimes_arrest_assault <-
  tidy_chicago_crimes_arrest[grepl("assault",
                                   tidy_chicago_crimes_arrest$Description,
                                   ignore.case = TRUE) |
                               grepl("assault",
                                     tidy_chicago_crimes_arrest$Primary_Type,
                                     ignore.case = TRUE), ]
# Any arrests dealing with theft?
tidy_chicago_crimes_arrest_theft <-
  tidy_chicago_crimes_arrest[grepl("theft",
                                   tidy_chicago_crimes_arrest$Description,
                                   ignore.case = TRUE) |
                               grepl("theft",
                                     tidy_chicago_crimes_arrest$Primary_Type,
                                     ignore.case = TRUE), ]

dim(tidy_chicago_crimes_arrest_theft)
dim(tidy_chicago_crimes_arrest_assault)
```

#### As we can see 

#### Assuming the description and primary type of each crime only have those two

#### strings that are about 207k thefts and about 93k assaults in Chicago reported

#### since 2001.