---
title: 'Week 10: Markov Chains & Random Walks'
author: "Jonathan Hernandez"
date: "October 30, 2018"
output:
  html_document: default
---

    18. Assume that a student going to a certain four-year medical school in northern
    New England has, each year, a probability q of flunking out, a probability r of
    having to repeat the year, and a probability p of moving on to the next year
    (in the fourth year, moving on means graduating).

    (a) Form a transition matrix for this process taking as states F, 1, 2, 3, 4, and
    G where F stands for flunking out and G for graduating, and the other states represent
    the year of the study
    
Answer: the matrix can be represented as the below. We got two absorbing states that is

the Flunking state and the Graduating state which are states that it is impossible to leave

(Flunk and you can't go back and Graduate and you don't come back)
```{r}
# Transition matrix P^1
P <- matrix(c(1, 0, 0, 0, 0, 0, 
              "q", "r", "p", 0, 0, 0, 
              "q", 0, "r", "p", 0, 0,
              "q", 0, 0, "r", "p", 0,
              "q", 0, 0, 0, "r", "p",
              0, 0, 0, 0, 0, 1), nrow = 6, ncol = 6, byrow = TRUE)
P
```

    (b) For the case q = 0.1, r = 0.2, p = 0.7 find the time a beginning student can
    expect to be in the second year. How long should this student expect to be in the medical
    school?
    
Answer: How long the student can expect to be in the second year is by looking at the matrix

$N = (I -Q)^{-1}$ and $n_{ij}$ gives the expected number of times that the process in the transient state if

it started in state $s_i$

There are two absorbing states and 4 transient states, let r = # of absorbing states and t = # transient states.

$I = 4 x 4$ identity matrix and 

Q is columns 1 through 4 and rows 1 =4 that is

```{r}
p <- 0.7
q <- 0.1
r <- 0.2

# replace strings with their respective probabilities
P <- gsub("p", p, P)
P <- gsub("q", q, P)
P <- gsub("r", r, P)
P <- mapply(P, FUN = as.numeric)
P <- matrix(data=P, ncol = 6, nrow = 6, byrow = FALSE)
rownames(P) <- c("F", "1", "2", "3", "4", "G")
colnames(P) <- c("F", "1", "2", "3", "4", "G")
P
# Compute Q
Q <- P[c("1", "2", "3", "4"), c("1", "2", "3", "4")]
Q

# Compute I
I <- diag(nrow = nrow(Q))
I

# Compute N
library(Matrix)
N <- solve(I - Q)
N
```

for state $s_1$ that is the first year and looking at state $s_2$ the 2nd year,

the 1st row and 2 column it is expected to take the student 1.09 times or 1.09 years to get

to the second year.

For the expected time the student will be in medical school we need to use the formula

$t = Nc$ where t is a column of vector whose i'th entry is $t_i$ and the expected number of steps before the chain

is absorbed. c is the nrow(N) x 1 column vector whose entries are all 1.

```{r}
c_vector <- rep(1, nrow(N))
t <- N %*% c_vector
t
```

Looking at the first entry, a beginning student should expect to graduate in about 4.13 years,

2nd year will take about 3.3 years, 3rd year will take about 2.34 years and in the final year will take about 1.25 years.

    (c) Find the probability that this beginning student will graduate.

Answer: we compute the Absorption probability $B = NR$ where $b_{ij}$ is the probability that an absorbing chain

will be absorbed in the absorbing state $s_j$ when it starts in $s_1$.

R is the rows of states 1 through 4 and columns of F and G in the orginal matrix P(absorbing states)

```{r}
R <- P[c("1", "2", "3", "4"), c("F", "G")]
R
B <- N %*% R
B
```

By looking at the first row that is the 1st year of the beginning student and the 2nd column,

we can see that the probability of this beginning student will graduate will be about 58.6%