---
title: "Chess Tournament Data Parsing"
author: "Jonathan Hernandez"
date: "September 19, 2016"
output: html_document
---

## This Assignment shows an example of given a raw file, using regular expressions

## extract the necessary data, and once clean and tidy, export to a CSV file for

## possible usage in a database. In this case it is working with chess Tournament

## Data and the goal is to scrape and clean the tournament data to extract the

## Player's name, their state, their pre-rating, total points accumated and the

## average pre-chess rating of player's opponents.


#### Load the data and make the columns more meaningful.

#### Note: I use a Ubuntu Linux environment to store my work so do change the 

#### directory of your choice on your OS.

```{r loadChessTournamentDataset, echo = TRUE, cache = TRUE}
# library(utils)
chess_data <- read.fwf("/home/jonboy1987/Desktop/CUNYSPS/IS607/Assignments/Project1/tournamentinfo.txt"
                       ,skip = 4, widths = c(5,3,32,6,6,6,6,6,6,6,6))

# give the chess_data dataset more better attribute names
colnames(chess_data) <- c("playerID_state","V2","PlayerName_prating","total_pts"
                          ,"round1","round2","round3","round4","round5","round6"
                          ,"round7")

head(chess_data)
```

#### We can extract the State using a regular expression by using column 1 of the

#### Chess data set by using "[A-Z]{2}" to get the states

```{r getStates, echo = TRUE}
# use grep and value = TRUE to show the strings that match the regexp
states <- grep("[A-Z]{2}", chess_data$playerID_state, value = TRUE)

# strip out the white spaces and store the Data
states <- trimws(states)
```

#### We will also extract the ID of each player; this will come in handy in 

#### computing the average pre-rating of opponents.

```{r getPlayerID, echo = TRUE}
ID <- grep("[0-9]+", chess_data$playerID_state, value = TRUE)
ID <- trimws(ID)
ID
```

#### The 2nd column in the chess_data set contains the name and Pre-Rating

#### extract.

```{r getPlayerName, echo = TRUE}
# "^[A-Z]+" matches all strings that start with uppercase letters
names <- grep("^[A-Z]+", chess_data$PlayerName_prating, value = TRUE)
names <- trimws(names)
```

```{r getPreRating, echo = TRUE}
# Extract the portion that starts with "R:"
pre_rating <- regmatches(chess_data$PlayerName_prating,
                         regexpr("R:\\s{1,2}[0-9]{1,4}",
                                 chess_data$PlayerName_prating))

pre_rating <- gsub("R:", "", pre_rating) # Remove the R:
pre_rating <- trimws(pre_rating) # remove whitespaces
```

```{r getTotalPoints, echo = TRUE}
# remove the "|" character from the total points field
chess_data$total_pts <- gsub('\\|', '', chess_data$total_pts)

total_points <- grep("^[0-9]+\\.[0-9]+",chess_data$total_pts, value = TRUE)
total_points <- trimws(total_points)
```

#### For the hardest part; to compute the average pre chess rating of opponents

#### First we will want to grab all the columns that contains all 7 rounds of 

#### the tournament and extract the player ID's from them.

```{r getRoundsDataSet, echo = TRUE}
rounds <- chess_data[,c("round1", "round2", "round3", "round4", "round5",
                        "round6", "round7")]

# Begin stripping down the rounds dataset
rounds <- rounds[-c(seq(from = 3, to = 192, by = 3)), ] # remove the -'s

# Remove the rows below the player ID's
rounds <- rounds[-c(seq(from = 2, to = 191, by = 2)), ]

# Add NA's to the entries that are empty 
# (player did not vs opponent for whatever reason)
rounds <- apply(rounds, 2, function(x) gsub("\\s{4}", NA, x))

# Now strip rest of characters so only the ID's and NA remain
rounds <- apply(rounds,2, function(x) gsub("^\\|[A-Z]\\s+", "",x))
rounds <- apply(rounds,2,as.numeric) # make the ID's numberic for indexing
head(rounds,15)
```

#### With the rounds dataset now cleaned to just ID's and NA's we can map

#### the ID to the pre-ratings

```{r IDtoRating, echo = TRUE}
# make a key-value pair so we can map 
ID_rating <- list(ID, pre_rating)
rounds <- apply(rounds, 1:2, function(x) ID_rating[[2]][x])
rounds <- apply(rounds,2,as.numeric) # make numeric for computing the average
rounds
```

#### Compute the Average Pre-Rating of opponents of each player

```{r get_avg_prerating, echo = TRUE}
avg_prerating <- round(apply(rounds, 1, mean, na.rm = TRUE), 0)
avg_prerating
```

```{r combineData, echo = TRUE}
chess_data_frame <- data.frame(name = names, state = states,
                               TotalPoints = total_points, 
                               Pre_Rating = pre_rating,
                               Average_Pre_Rating = avg_prerating)
chess_data_frame

# Finally export the new Clean dataset into a .csv file
write.csv(chess_data_frame,
          "/home/jonboy1987/Desktop/CUNYSPS/IS607/Assignments/Project1/chess_tournament_data.csv",
          row.names = FALSE)
```