---
title: 'Week 12: Regression Analysis in R: Multiple Linear Regression'
author: "Jonathan Hernandez"
date: "November 12, 2018"
output:
  html_document: default
---

    Using R, build a multiple regression model for data that interests you.
    Include in this model at least one quadratic term,bone dichotomous term, 
    and one dichotomous vs. quantitative interaction term. 
    Interpret all coefficients. Conduct residual analysis. Was the linear model appropriate? Why or why not?
    
- For this discussion, I will look at the Kaggle Dataset "Medical Cost Personal Datasets"
link can be found here to download: https://www.kaggle.com/mirichoi0218/insurance

- This dataset looks at medical insurance costs charges for various people based on several
factors like number of children, region of residency, age etc.

- I will make a multiple linear regression model and make a best-fit line for computing medical costs.

- Start by loading the data and viewing attributes

```{r}
insurance <- read.csv("insurance.csv")
dim(insurance)
summary(insurance)
str(insurance)
head(insurance, n = 10)
```

- Dataset is well clean and tidy; nothing else to do to it.
Let's plot some graphs.

```{r}
par(mfrow=c(1,2))
hist(insurance$bmi, xlab = "BMI (Body Mass Index)",
     main = "Histogram of BMI")
hist(insurance$charges, xlab = "Medical Charges",
     main = "Histogram of Medical Charges") # looks more right-skewed

# let's do some plotting of charges against other factors like sex and smoker and region
par(mfrow=c(1,3))
with(insurance, plot(charges ~ smoker + sex + region))
```

- We see that BMI is nearly normally distributed, medical charges is right-skewed and
there are many outliers for high medical charges against both genders and various regions.

- We also see that the median is about the same for all regions, and genders.

- Note that for smokers, medical charges are much higher than normal ones which we
should expect.

- Now, let's fit a multiple regression model, let have the explanatory variables as

    - sex (categorical)
    - bmi (numerical, continous)
    - age (numerical, discrete)
    - smoker (categorical)
    - charges (numerical, continous)
    
- Let's make a multiple regression model of the following equation:

$$
\begin{aligned}
\widehat{charges} = \beta_0 + \beta_1 * Sex + \beta_2 * bmi + 
\beta_3 * age + \beta_4 * smoker + \beta_5 (bmi*sex)
\end{aligned}
$$
```{r}
lm_insurance <- lm(charges ~ sex + bmi + age + smoker + bmi*sex, data = insurance)
summary(lm_insurance)
par(mfrow=c(2,2))
hist(lm_insurance$residuals, main = "Histogram of Residuals", xlab= "")
plot(lm_insurance$residuals, fitted(lm_insurance))
qqnorm(lm_insurance$residuals)
qqline(lm_insurance$residuals)
```

- We see that the residuals histogram is somewhat normal but the residuals vs fitted
values doesn't show constant variance which is not good for a multiple regression model.


- The equation of this multiple regression model is as follows:

$$
\begin{aligned}
\widehat{charges} = -11717.37 + 54.1*sex + 325.78*bmi + 259.47*age + 23836.07*smoker - 5.33(bmi*sex)
\end{aligned}
$$

- Note the variables

    - sex = 1 for male and 0 for female
    - smoker = 1 for male and 0 for female
    
- What does this tell us? Let's look at the details of the summary in more detail

    - Coefficients:
        - Intercept: This tells us that leaving all other terms constant, on average
        the estimated medical charge is about $-11717.36 which logically won't make sense
        and is good there are other terms in the model.
        - Sex: If a person is male and leaving all other terms constant, he can expect
        to pay about $54.1 in medical costs.
        - BMI: Leaving all other terms constant, a person can be expected to pay about
        $325.78 in medical charges per BMI value.
        - Age: Leaving all other terms constant, a person can be expected to pay about
        $259.47 in medical expenses multiplied by their age (A 31 year old will pay about $8043.57)
        - Smoker: A person who smokes and leaving all variables constant can expect to pay $23836.07
        - Sex*BMI: A male can expect to pay holding all other variables constant can expect to pay
        $-5.326 which doesn't make sense logically.
    - P-values of coefficients:
        - The p-values of the intercept, bmi, age and male smokers are very low and we can reject the 
        null hypothesis (H_0 = 0) and favor the alternative (H_A != 0) that is the true coefficients is not 0
        - For Males and Male*bmi, we fail to reject the null hypothesis and thus these coefficients are very
        close to 0 and can be excluded in our model.
    - Residual Standard Error: The residual standard error of 6097 is the standard deviation and is a bit far
    from the good fit of points.
    - R-squared/Adjusted R^2: values of 0.7475 and 0.7466 respectively, this means that about 75% of the data
    fall into the regression line.
    - F-statistic: value of 788.6 with a small p-value < 2.2e-16 means that the features selected are better
    than the intercept-only model which as described before makes sense as a intercept only model gives a negative
    medical cost which doesn't apply or make sense.

- Let's predict the medical cost of myself:

    - I'm 31 years old: Age = 31
    - I don't smoke: smoke = 0
    - I'm male: male = 1
    - My BMI is: 31.3
  
```{r}
my_predicted_medical_cost <- predict(lm_insurance,
                                     data.frame(age=31, smoker="no", 
                                                bmi=31.3, sex="male"))
my_predicted_medical_cost
```

- So the multiple regression model that I made says my approximate medical costs is 
about $6410.44.

- So far this year on a personal note, I've only spent about $350 in medical costs,
which my predicted model is no where near close and overestimates my medical costs!

- I would not be comfortable with using the model above as there are coefficients that can be removed or probably
added for better accuracy and properly modeling and predicting medical costs. The residual standard error as well
as the Q-Q plots show that the model is not a good fit for the data. One good thing I can say about the model
is that the BMI and Age coefficients make sense as the more your BMI is and older, you are more likley to have 
more health problems and have more medical costs to pay.

- Future work can be done to add more coefficients, transforms and possibly use non-linear regression to better predict 
medical costs.
