---
title: "Week 13 Univariate & Multivariate Calculus"
author: "Jonathan Hernandez"
date: "November 4, 2018"
output: html_document
---

    A standard soda can is roughly cylindrical and holds 355cm^3
    of liquid. What dimensions should the cylinder be to min-
    imize the material needed to produce the can? Based on
    your dimensions, determine whether or not the standard
    can is produced to minimize the material costs.
    
- We are given the volume of Soda can and the formula for the volume of the cylinder is

$$
\begin{aligned}
V = \pi r^2h = 355cm^3
\end{aligned}
$$

- This is a optimization problem and we are trying to minimize the Surface Area

(material needed) and find the radius r and height h to make the soda can given the

volume of the standard soda can.

- Surface Area of a Cylinder = Area of circular ends + curved surface area
that is

$$
\begin{aligned}
A = 2\pi r^2 + 2\pi rh
\end{aligned}
$$

- This is a equation of two unknowns but we can solve for h with the information given

and then plug it into the equation for A

$h = 355/(\pi r^2)$

- Putting it back into A gives

$$
\begin{aligned}
A = 2\pi r^2 + 2\pi r(355/(\pi r^2)) \\
= 2\pi r^2 + 710/r
\end{aligned}
$$
- Now we have A as a function of r that is $A(r)$

- To find the minima or critical points we compute the derivative of A with respect

to r and set it to 0

to find r such that the Area is minimal:

$$
\begin{aligned}
A(r) = 2\pi r^2 + 710/r \\
A^{'}(r) = 4\pi r - 710/r^2 \\
\text{set A'(r) = 0} \\
0 = 4\pi r - 710/r^2 \\
710/r^2 = 4\pi r \\
r = \sqrt[3]{(710/4\pi)}
\end{aligned}
$$

- With r found, we can go back to the given volume to solve for h and then finally

comupte A(r,h) to find the minimum surface area or material.

```{r}
# Compute the surface area of a cylinder with given radius r and height h
A <- function(r,h){
  surfacearea <- (2*pi*r^2) + (2*pi*r*h)
  return(surfacearea)
}

# value of r that minimizes A
r_min_A <- (710/(4*pi))^(1/3)
r_min_A

# value of h that minimizes A given r
h_min_A <- 355/(pi*(r_min_A^2))
h_min_A

# plug in the values or r and h that minimize the surface area A
min_A <- A(r_min_A, h_min_A)
min_A
```

- We see that by having the $r = 3.83cm$, $h = 7.67cm$ we get the minimum material used

is about $277.54cm^2$. To determine whether or not the standard can is produced to minimize the 
material costs we compute $V=\pi r^2h$ with the given r and h.

```{r}
V <- (pi*r_min_A^2)*h_min_A
V
```

- We see the minimum material is produced to make the standard soda can.